---
title: "Chapter 18 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(nycflights13)
library(hms)
```

### 18.2.5

---

1.

```{r}

ymd(c("2010-10-10", "bananas"))

```

2.

`tzone` specifies the time zone, which is by default set as your computer's.

3.

```{r}

tibble(
d1 <- "January 1, 2010",
readr::parse_date(d1, "%B %d, %Y"),
lubridate::parse_date_time(d1, "%B %d, %Y")
)

tibble(
d2 <- "2015-Mar-07",
readr::parse_date(d2, "%Y-%b-%d"),
lubridate::ymd(d2)
)

tibble(
d3 <- "06-Jun-2017",
readr::parse_date(d3, "%d-%b-%Y"),
lubridate::parse_date_time(d3, "%d-%b-%Y")
)

tibble(
d4 <- c("August 19 (2015)", "July 1 (2015)"),
readr::parse_date(d4, "%B %d (%Y)"),
lubridate::parse_date_time(d4, "%B %d (%Y)")
)

tibble(
d5 <- "12/30/14", # Dec 30, 2014
readr::parse_date(d5,"%m/%d/%y"),
lubridate::mdy(d5)
)

tibble(
t1 <- "1705",
lubridate::parse_date_time(t1, "%H%M") |> hms::as_hms(),
readr::parse_time(t1, "%H%M")
)

tibble(
t2 <- "11:15:10.12 PM",
readr::parse_time(t2, "%I:%M:%OS %p"),
lubridate::parse_date_time(t2, "%I:%M:%OS %p")|> hms::as_hms()
)


```

### 18.3.4

---

1.

Looking at the distribution of departure times within a given month reveals no 
clear changes across the year: there is a peak of departure early in the morning 
and around dinner time.

```{r}
#Create a new data frame containing shceduled departure date, month, and time
flights_dept_data <- flights |> 
  mutate(
    departure_month = lubridate::month(month, label = TRUE, abbr = FALSE, locale = "en_GB"),
    departure = make_datetime(year, month, day, hour, minute),
    departure_scheduled = hms::as_hms(departure)
  )
  
#Plot the freqplot of scheduled departure time within a given month
flights_dept_data |> 
  ggplot(aes(x = departure_scheduled)) +
  geom_freqpoly() +
  facet_wrap(~ departure_month)

```

2.

```{r}

# Add extra information about departure time, reported and actual delay, and any 
# difference
flights_dept_data <- flights_dept_data |> 
  mutate(
    departure_actual = hms(hours = dep_time %/% 100, minutes = dep_time %% 100),
    departure_delay_reported = hms(minutes = dep_delay),
    departure_delay_actual = hms(seconds_to_period(departure_actual - departure_scheduled)),
    departure_delta = hms(seconds_to_period(departure_delay_actual - departure_delay_reported))
  ) 

# Filter out observations where the reported and actual delay don't match
flights_dept_data |> 
  filter(departure_delay_reported != departure_delay_actual)
  
# These mistakes seems to be due to the fact that some flights are delayed until
# the next day.
flights_dept_data |> 
  select(dep_time, sched_dep_time, contains(c("departure_delay"))) |> 
  filter(departure_delay_reported != departure_delay_actual)

# Let's correct these observations specifically, by instead accounting for that
# these delays were into the next day 

flights_dept_data_corrected <- flights_dept_data |> 
  mutate(
    departure_delay_actual = if_else(
      departure_delay_reported != departure_delay_actual,
      hms(seconds_to_period(24 * 60 * 60 - (departure_scheduled - departure_actual))),
      departure_delay_actual
                              ),
    departure_delta = hms(seconds_to_period(departure_delay_actual - departure_delay_reported))
  )

#Now all reported and actual times seem to overlap
flights_dept_data_corrected |> 
  filter(departure_delay_reported != departure_delay_actual)


```

3.

Calculating and comparing `actual_air_time` to reported `air_time` shows that the
latter is including time zone difference.

```{r}

flights_dept_data_corrected |> 
  mutate(
    arrival_actual = hms(hours = arr_time %/% 100, minutes = arr_time %% 100),
    air_time = hms(hours = air_time %/% 100, minutes = air_time %% 100),
    actual_air_time = hms(seconds_to_period(arrival_actual - departure_actual))
  ) |> 
  select(departure_actual, arrival_actual, air_time, actual_air_time)


```

4.

If we plot the average delay by time of day, we see that using the actual 
departure time inflates the average delay of the early (night) flights - this is
probably due the actual departure includes a lot of delayed flights from the day 
before. As such, we should use the scheduled departure time.

```{r}

#Filter out the variables of interest
filtered <- flights_dept_data_corrected |> 
  select(departure_scheduled, departure_actual, departure_delay_actual)

#Calculate the average and plot, for scheduled or actual departure time
filtered |> 
  pivot_longer(1:2, values_to = "time") |> #Pivot into one column before summarise
  group_by(name, time) |> 
  summarise(
    avg_delay = hms(seconds_to_period(mean(departure_delay_actual, na.rm = TRUE))),
    n = n(),
    .groups = "drop"
  ) |> 
  ggplot(aes(x = time, y = avg_delay)) +
  geom_col() +
  facet_wrap(~ name)

```

5.

Not correcting for any other factors (such as airport, destination, etc): go on
a Saturday.

```{r}

wday_delay <- flights_dept_data_corrected |> 
  mutate(wday = wday(make_date(year,month,day), label = TRUE, abbr = FALSE, locale = "en_GB")) |> 
  select(wday,departure_delay_actual)

wday_delay |> 
  group_by(wday) |> 
  summarise(
   avg_delay = hms(seconds_to_period(mean(departure_delay_actual, na.rm = TRUE))),
   n = n()
  ) |> 
  ggplot(aes(x = wday, y = avg_delay)) +
  geom_col()

```

6.

They are both not really continuous:

```{r}

diamonds_plot <- ggplot(diamonds, aes(carat)) +
  geom_histogram(bins = 500)

flights_plot <- ggplot(flights, aes(sched_dep_time)) +
  geom_histogram(bins = 250)

library(patchwork)
diamonds_plot + flights_plot

```


