---
title: "Chapter 10 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)
```

### 10.2.1

---

1.

```{r}
mpg |> 
  ggplot() + 
  geom_point(aes(x = hwy, y = displ), shape = 24, fill = "pink")
```

2.

The `color = "blue"` needs to be moved out of aes when it is set to a value 
instead of mapped to a variable.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, color = "blue"))

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy), color = "blue")
```

3.

Stroke sets the width of the border (if your shape has a border).

4.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, color = displ < 5))
```

### 10.3.1

---

1. 

`geom_line()` for a line, `geom_boxplot()` for a boxplot, `geom_area()` for an 
area plot.

2.

`show.legend = FALSE` hides the geom from the legend. This was used previously
to maintain the regular plot shape to ease comparison (otherwise legend shrinks
the plot area).

3.

`se` plots the confidence interval of `geom_smooth()`.

4.

Going per row, from left to right:

```{r}
mpg |> #1
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(se = FALSE)

mpg |> #2
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(aes(group = drv), se = FALSE)

mpg |> #3
  ggplot(aes(x = displ, y = hwy, group = drv, color = drv)) + 
  geom_point() +
  geom_smooth(se = FALSE)

mpg |> #4
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point(aes(group = drv, color = drv)) +
  geom_smooth(se = FALSE)

mpg |> #5
  ggplot(aes(x = displ, y = hwy, group = drv, linetype = drv)) + 
  geom_point(aes(color = drv)) +
  geom_smooth(se = FALSE)

mpg |> #6
  ggplot(aes(x = displ, y = hwy, group = drv, linetype = drv)) + 
  geom_point(aes(fill = drv), shape = 21, color = "white", stroke = 2 )

```

### 10.4.1

---

1. 

```{r}

mpg |> 
  ggplot(aes(x = hwy, y = displ)) +
  geom_point() +
  facet_wrap(~cty)

```

2.

It means that the specific `drv` and `cyl` combination is not existing in the 
data.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = drv, y = cyl))
```

3.

It controls whether the variable will be displayed side-by-side or stacked.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4.

For larger data sets, color gets busy and, unless specified, is only recommeded 
for six separate group members (colors).

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

5.

`ncol`or `nrow` controls `facet_wrap()` plot layout, and is inteded for use with
a single variable with many levels. Instead, `facet_grid()` is intended for 
facetting two separate variables against eachother in a matrix layout.

6.

It is much easier to compare when they are stacked.

```{r}
ggplot(mpg, aes(x = displ)) + 
  geom_histogram() + 
  facet_grid(drv ~ .)

ggplot(mpg, aes(x = displ)) + 
  geom_histogram() +
  facet_grid(. ~ drv)
```

7.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~drv, ncol = 1, strip.position = "right")
```

### 10.5.1

---

1.

