---
title: "Chapter 2 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)  
library(palmerpenguins)  
library(ggthemes)
```




### 2.2.5

---

1.

```{r}
penguins %>% nrow()
```

2.

```{r, eval=FALSE}
?penguins
```

3.  

```{r}
ggplot(data = penguins, 
       mapping = aes(
         x = bill_length_mm, 
         y = bill_depth_mm ) 
       )+ 
  geom_point()

```


4.  

Consider using a box plot instead.

```{r}
ggplot( data = penguins, 
        mapping = aes(
          x = species, 
          y = bill_depth_mm)
        )+ 
  geom_point() 
```



5.

You need to define aesthetics.


```{r, error = TRUE}
ggplot(data = penguins) + 
  geom_point()
```
  

6.

`na.rm` removes the observations (rows) where an NA is noted (i.e.
missing data).


```{r}
ggplot(data = penguins, 
       mapping = aes(
         x = bill_length_mm,
         y = bill_depth_mm)
       )+
  geom_point(na.rm = T)
```

7.

```{r}
ggplot( data = penguins, 
        mapping = aes(
          x = bill_length_mm, 
          y = bill_depth_mm)
          )+ 
  geom_point(na.rm = T)+ 
  labs(caption = "Data come from the palmerpenguins package.")
```



8.

```{r}
ggplot( data = penguins, 
        mapping = aes(
          x = flipper_length_mm,
          y = body_mass_g, 
          color = bill_depth_mm) 
        )+ 
  geom_point()+ 
  geom_smooth()
```

9.

The plot will look like 8\, but with curves and geom_point color by
island\, and no confidence interval. 

```{r}
ggplot( data = penguins,
        mapping = aes(
          x = flipper_length_mm,
          y = body_mass_g, 
          color = island) 
        )+ 
  geom_point()+ 
  geom_smooth(se = FALSE)

```

10.

They two plots will look the same as the same aesthetics are called, in different ways. 

```{r}
ggplot(data = penguins, 
        mapping = aes(
          x = flipper_length_mm, 
          y = body_mass_g)
        )+ 
  geom_point()+ 
  geom_smooth()

ggplot()+
  geom_point(data = penguins, 
             mapping = aes(
               x = flipper_length_mm, 
               y = body_mass_g)
             )+ 
  geom_smooth(data = penguins,
              mapping = aes(
                x = flipper_length_mm, 
                y = body_mass_g)
              )
```

### 2.4.3

---

1.


```{r}
ggplot(penguins, 
       aes(y = species)
       )+ 
  geom_bar()
```

2. 
Fill is more useful for bars.

ggplot(penguins, aes(x = species)) +
geom_bar(color = "red")
```{r}
ggplot(penguins, aes(x = species)) + geom_bar(fill = "red")

ggplot(penguins, aes(x = species)) + geom_bar(color = "red")
```



3.

`bins` overrides `binwidth` by defining a set number of bins that the
data is divided into.

4.

```{r}
diamonds %>% summarise(min = min(carat), max = max(carat))
```

```{r, warning=FALSE, message=FALSE}
diamonds %>% 
  ggplot(aes(x = carat))+ 
  geom_histogram()+ 
  labs(title = "undefined")

diamonds %>% 
  ggplot(aes(x = carat))+ 
  geom_histogram(binwidth = 0.01)+
  labs(title = "0.01")

diamonds %>% 
  ggplot(aes(x = carat))+ 
  geom_histogram(binwidth = 0.1)+
  labs(title = "0.1")

diamonds %>%
  ggplot(aes(x = carat))+ 
  geom_histogram(binwidth = 0.2)+
  labs(title = "0.2")

diamonds %>%
  ggplot(aes(x = carat))+ 
  geom_histogram(binwidth = 0.5)+
  labs(title = "0.5")

diamonds %>%
  ggplot(aes(x = carat))+ 
  geom_histogram(binwidth = 1)+
  labs(title = "1")
```

### 2.5.5

---

1. 

Easiest is to count number of character (categorical) versus numeric
(dbl and int)  

```{r}
mpg %>% glimpse()
numeric <- mpg %>% select(where(is.numeric)) %>% length()
cat <- mpg %>% select(!where(is.numeric)) %>% length()
numeric
cat
```

2.

Shape cannot be mapped to a continuous variable.

```{r}
mpg %>% 
  ggplot(aes(x = hwy, y = displ))+ geom_point()

mpg %>% 
  ggplot(aes(x = hwy, y = displ, color = year))+ geom_point()

mpg %>% 
  ggplot(aes(x = hwy, y = displ, size = year))+ geom_point()

mpg %>% 
  ggplot(aes(x = hwy, y = displ, color = year, size = year))+ geom_point()

```


3.

```{r}
mpg %>% 
  ggplot(aes(x = hwy, 
             y = displ, 
             color = year, 
             size = year,
             linewidth = cty)
         )+ 
  geom_point()
```

4.

```{r}
mpg %>% 
  ggplot(aes(
    x = year, 
    y = year, 
    color = year, 
    size = year,
    linewidth = year)
    )+ 
  geom_point()
```

5.

```{r}
penguins %>% 
  ggplot(aes(
    x = bill_depth_mm, 
    y = bill_length_mm, 
    color = species)
    )+ 
  geom_point()

penguins %>% 
  ggplot(aes(
    x = bill_depth_mm, 
    y = bill_length_mm, 
    color = species)
    )+ 
  geom_point()+ 
  facet_wrap(~species)
```

6.

```{r}
ggplot(data = penguins, 
       mapping = aes(
         x = bill_length_mm, 
         y = bill_depth_mm, 
         color = species, 
         shape = species)
       )+ 
  geom_point()+
labs(color = "Species")

ggplot( data = penguins, 
        mapping = aes(x = bill_length_mm,
                      y = bill_depth_mm, 
                      color = species, 
                      shape = species) 
        )+ 
  geom_point()+
labs(color = "Species", shape = "Species")
```

7. 
They show island or species on x, and fill by the other variable -
thereby comparing island or species, depending on what is on X.

```{r}
ggplot(penguins, 
       aes(
         x = island, 
         fill = species)
       )+ 
  geom_bar(position = "fill") 

ggplot(penguins, 
       aes(
         x = species, 
         fill = island)
       )+
  geom_bar(position = "fill")
```

### 2.6.1

---


1.

In `ggsave`, the `plot` option defaults to `last_plot()`, unless otherwise specified.

2. 

You change file ending to .pdf, you can check available output formats
using `?ggsave()`.
