---
title: "Chapter 16 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(babynames)
library(ggrepel)
```

### 16.3.5

---

1.


```{r}

babynames_vowels <- babynames |> 
  distinct(name) |> 
  mutate(
    n_vowels = str_count(name, "[AEIOUYaeiouy]"),
    n_total = str_length(name),
    prop_vowel = n_vowels / n_total
  )

babynames_vowels |> 
  arrange(desc(n_vowels)) |>  #Mariaguadalupe has the most total vowels
  arrange(desc(prop_vowel)) |> #Many names are 100% vowels
  count(prop_vowel == 1) #40 to be exact

```

2.

```{r, error=TRUE}

c("a/b/c/d/e") |> 
  str_replace_all("/", "\\") |> 
  str_replace_all("\\", "/")

```

3.

If I understand the question, you can use `str_replace_all(ignore.case = TRUE)`.

4.

```{r}

a_swedish_phone_number <- c("0701112233", "+46701112233")
str_extract(a_swedish_phone_number, "\\+?[0-9]*")
```

### 16.4.7

---

1.

```{r}

r"{"'\}"
r"{"$^$"}"

```

2.

To create a regex matching a literal `\`, we need to use `"\\\\"`:

  - `"\"` escapes the second ", resulting in an "unclosed" (open) character string.
  - `"\\"` will not match a literal \\, because it results in a single \\, which 
  has a function - and will not be interpreted as a literal \\.
  - `"\\\"` runs into a similar issue as the first line: the first \\ will 
  escape the second which escapes the third and creates an open character string.
  
3.

```{r}

str_view(words, "^y") #a
str_view(words, "^[^y]") #b
str_view(words, "x$") #c
str_view(words, "^\\w{3}$") #d
str_view(words, "^\\w{7,}$") #e
str_view(words, "([AEIOUYaeiouy][^AEIOUYaeiouy])") #f
str_view(words, "([AEIOUYaeiouy][^AEIOUYaeiouy]){2,}") #g
str_view(words, "^([AEIOUYaeiouy][^AEIOUYaeiouy]){1,}$") #h

```

4.

```{r}

str_4 <- c(
  "airplane", "aeroplane", 
  "aluminum", "aluminium", 
  "analog", "analogue", 
  "ass", "arse", 
  "center", "centre", 
  "defense", "defence", 
  "donut", "doughnut", 
  "gray", "grey", 
  "modeling", "modelling", 
  "skeptic", "sceptic", 
  "summarize", "summarise"
          )

str_view(str_4, "a(i|e)ro?plane") #1
str_view(str_4, "alumi(ni|n)um") #2
str_view(str_4, "analog(ue)?") #3
str_view(str_4, "^a(ss|rse)$") #4
str_view(str_4, "cent(er|re)") #5
str_view(str_4, "defen(s|c)e") #6
str_view(str_4, "do(ugh)?nut") #7
str_view(str_4, "gr(a|e)y") #8
str_view(str_4, "model+ing") #9
str_view(str_4, "s(k|c)eptic") #10
str_view(str_4, "summari(z|s)e") #11

```

5.

```{r}

words_scramble <- words |> 
str_replace("([a-zA-Z])(.*)([a-zA-Z])","\\3\\2\\1")

as_tibble(words) |> 
  filter(words %in% words_scramble) |> 
  print(n = Inf)


```

6.

  a. `^.*$` captures a whole word
  b. `"\\{.+\\}"` captures a string that contains any number of character
  enclosed by {}
  c. `\d{4}-\d{2}-\d{2}` will match a date format YYYY-MM-DD
  d. Will match "\\{4}"
  e. Will match strings with three repeats of .. separated by a single \\
  f. Will match three repeats of the first character
  g. Will match two repeats
  
### 16.6.4

---

1.

```{r}

#a
str_view(words, "^x|x$")
words[
  str_detect(words, "^x") |
  str_detect(words, "x$")
]

#b
str_view(words, "^[aieouy].*[^aieouy]$")

words[
  str_detect(words, "^[aieouy]") &
  str_detect(words, "[^aieouy]$")
]

#c
str_view(words, "a.*e.*i.*o.*u")

words[
  str_detect(words, "a") &
    str_detect(words, "e") &
    str_detect(words, "i") &
    str_detect(words, "o") &
    str_detect(words, "u")
]

```

2.


