---
title: "Chapter 20 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(nycflights13)
library(hms)
```

### 20.2.4

---

1.

Weather and airports share the primary keys `origin` or `faa` (respectively).

2.

If weather it extended `origin`, it wouldn't change any of its keys - only extend.

3.

```{r}

#Identify duplicates
weather_dups <- weather |> 
  count(year, month, day, hour, origin) |> 
  filter(n > 1)

# Find them in the full data set - look for additional information.
# A quick google tells me that this was the DST date, probably the cause for 
# duplicate values.
left_join(weather_dups, weather)


```

4.

We coud create a new data frame with a column stating the `holiday` name, and use
`year`, `month`, and `day` as primary keys.

5.

### 20.3.4

1.

```{r}

# Make a date from year, month, and day
flights_date <- flights |> 
  mutate(
    date = make_date(year, month, day)
  )

#Calculate time elapsed in hours from the minimum date
min_date <- pull(flights_date, date)[1]

flights_date <- flights_date |> 
  mutate(
    time_elapsed_hours = as.numeric(difftime(date, min_date, units = "hours"))
  )

# Create new variable group: dividing time into 48 hour bins
flights_date <- flights_date |> 
  mutate(
    group = floor(time_elapsed_hours / 48) + 1
  )

# Calculate the avgerage delay by 48 hour group
flights_date <- flights_date |> 
  mutate(avg_delay = mean(dep_delay, na.rm = TRUE), .by = group)

# Pull out top 5 worst 48 hour periods
flights_date_worst <- flights_date |> 
  arrange(desc(avg_delay)) |> 
  distinct(date, group, avg_delay) |> 
  slice(1:10) |> # 5 time periods * 2 days each
  mutate(delay = "worst")

# Pull out top 5 best 48 hour periods
flights_date_best <- flights_date |> 
  arrange(avg_delay) |> 
  distinct(date, group, avg_delay) |> 
  slice(1:10) |> # 5 time periods * 2 days each
  mutate(delay = "best")

# Bind their rows into a new tibble
flights_date_worst_best <- bind_rows(flights_date_worst, flights_date_best)

flights_date_worst_best |> 
  print(n = Inf)

# Add date, time elapsed, and group, to weather (as was done before)
weather_date <- weather |> 
  mutate(
    date = make_date(year, month, day),
    time_elapsed_hours = as.numeric(difftime(date, min_date, units = "hours")),
    group = floor(time_elapsed_hours / 48) + 1
  ) 

#Add in information to weather_date if day is among worst or best
weather_date <- weather_date |> 
  left_join(flights_date_worst_best)

#Plot a boxplot of weather data, color by worst or best (least) delay
weather_date |> 
  pivot_longer(temp:wind_gust, names_to = "parameters", values_to = "measurement") |> 
  filter(delay %in% c("worst","best")) |> 
  ggplot(aes(x = delay, y = measurement, color = delay)) +
  geom_boxplot() +
  facet_wrap(~ parameters, scales = "free_y")


```

2.

```{r}

# Provided code
flights2 <- flights |> 
  select(year, time_hour, origin, dest, tailnum, carrier)

top_dest <- flights2 |>
  count(dest, sort = TRUE) |>
  head(10)

# Pull out the flights to top destinations
top_dest_flights <- flights_date |> 
  filter(dest %in% pull(top_dest, dest)) |> 
  mutate(time = parse_date_time(sched_dep_time, "%H%M") |> as_hms()) |> 
  distinct(dest,origin, flight, date, time) |> 
  arrange(date, dest)

```

3.

```{r}


# Create a date and time variable for joining to top flights dataframe
weather_dttm <- weather |> 
  mutate(
    date = date(time_hour),
    time = as_hms(time_hour)
  ) |> 
  select(!c(year:hour, time_hour))

# Joining the two shows the problem with using an exact time - a lot of "missing" 
# weather data
top_dest_weather <-left_join(top_dest_flights, weather_dttm)

# Workaround - bin the data by time of day (top flights dataframe)
top_dest_flights_timeofday <- top_dest_flights |> 
  mutate(
    time_of_day = case_when(
      hour(time) <= 5 ~ "night",
      hour(time) > 5 & hour(time) <= 11 ~ "morning",
      hour(time) > 11 & hour(time) <= 17 ~ "mid day",
      hour(time) > 17 & hour(time) <= 23.59 ~ "evening",
                  )
  ) 

# Workaround - bin the data by time of day (weather dataframe)
weather_dttm_timeofday <- weather_dttm |> 
  mutate(
    time_of_day = case_when(
      hour(time) <= 5 ~ "night",
      hour(time) > 5 & hour(time) <= 11 ~ "morning",
      hour(time) > 11 & hour(time) <= 17 ~ "mid day",
      hour(time) > 17 & hour(time) <= 23.59 ~ "evening",
    )
  ) |> 
  select(!time) |> 
  group_by(origin, date, time_of_day) |> 
  mutate(
    across(
      temp:visib, ~ mean(.x, na.rm = TRUE)
    )
  ) |> 
  distinct(
    origin, date, time_of_day, .keep_all = TRUE
  )
  
# Now join by time_of_day instead of an exact time
top_dest_weather_time_of_day <- left_join(
  top_dest_flights_timeofday, weather_dttm_timeofday, 
  by = join_by("origin", "date", "time_of_day")
)

```

4.
from `?planes`:
*Plane metadata for all plane tailnumbers found in the FAA aircraft registry.
American Airways (AA) and Envoy Air (MQ) report fleet numbers rather than tail 
numbers so can't be matched.*

```{r}

flights |> 
  anti_join(planes, join_by("tailnum")) |> 
  summarise(
    n = n(),
    .by = carrier,
    .order = TRUE
  ) |> 
  mutate(
    prop = n / sum(n)
  )

```

From this, we see that ~90% of the missing tailnums are from these carriers.

5.

```{r}

# Isolate all distinct carrier and tailnum observations
all_carrier_tailnum <- flights |> 
  filter(!is.na(tailnum)) |> 
  select(tailnum, carrier) |> 
  distinct()

# OBS! These two following steps are optional but helpful to arrange and inspect
# if the code is producing the desired result

# Count the observations of tailnums (i.e. >2 is duplicated)
dups <- all_carrier_tailnum |> count(tailnum) |> arrange(desc(n))

# Join them
all_carrier_tailnum <- left_join(all_carrier_tailnum, dups) |>  arrange(desc(n), tailnum)

# Create a new variable "active_carriers", that combines all carriers for a
# given tailnum
all_carrier_tailnum <- all_carrier_tailnum |>
  mutate(active_carriers = str_c(carrier, collapse = "_"), .by = tailnum)


# Add this information to the planes dataframe
left_join(planes, all_carrier_tailnum, by = join_by(tailnum)) |> 
  relocate(active_carriers, before = 1) |> 
  arrange(desc(n))


```

