---
title: "Chapter 2 excercises"
author: "hugoakerstrand"  
output: github_document
---


```{r setup, include=FALSE}
# Ladda in dplyr globalt för alla kodblock
library(tidyverse)  
library(palmerpenguins)  
library(ggthemes)
```

# 2.2.5

**1.**
```{r}
penguins %>% nrow()
```

#2

?penguins

#3

ggplot(
    data = penguins,
    mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    )+
    geom_point()
    
#4

ggplot(
    data = penguins,
    mapping = aes(x = species, y = bill_depth_mm)
    )+
    geom_point()
#Consider a box plot

#5

#You need to define aesthetics

#6

#na.rm removes the observations (rows) where an NA is noted (i.e. missing data)
ggplot(
    data = penguins,
    mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    )+
    geom_point(na.rm = T)
    
#7

ggplot(
    data = penguins,
    mapping = aes(x = bill_length_mm, y = bill_depth_mm)
    )+
    geom_point(na.rm = T)+
    labs(caption = "Data come from the palmerpenguins package.")


#8

ggplot(
    data = penguins,
    mapping = aes(x = flipper_length_mm,
                  y = body_mass_g,
                  color = bill_depth_mm
                 )
    )+
geom_point()+
geom_smooth()

#9
#The plot will look like 8, but with curves and geom_point color by island
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)

#10
#They will look the same as the same aesthetics are used
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )

#2.4.3

#1
ggplot(penguins,
       aes(y = species)
       )+
geom_bar()

#2
#Fill is more useful for bars
ggplot(penguins, aes(x = species)) +
  geom_bar(color = "red")

ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "red")

#3
#bins overrides binwidth by defining a set number of bins that the data is divided into

#4
diamonds %>% summarise(min = min(carat), max = max(carat))

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram()+
labs(title = "undefined")

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram(binwidth = 0.01)+
labs(title = "0.01")

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram(binwidth = 0.1)+
labs(title = "0.1")

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram(binwidth = 0.2)+
labs(title = "0.2")

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram(binwidth = 0.5)+
labs(title = "0.5")

diamonds %>%
ggplot(aes(x = carat))+
geom_histogram(binwidth = 1)+
labs(title = "1")

#2.5.5

#1
#Easiest is to count number of character (categorical) versus numeric (dbl and int)
mpg %>% glimpse()

#2
mpg %>%
ggplot(aes(x = hwy, y = displ))+
geom_point()

mpg %>%
ggplot(aes(x = hwy, y = displ, color = year))+
geom_point()

mpg %>%
ggplot(aes(x = hwy, y = displ, size = year))+
geom_point()

mpg %>%
ggplot(aes(x = hwy, y = displ, color = year, size = year))+
geom_point()

#shape cannot be mapped to a continuous variable

#3

mpg %>%
ggplot(aes(x = hwy, y = displ, color = year, size = year, linewidth = cty))+
geom_point()

#4
mpg %>%
ggplot(aes(x = year, y = year, color = year, size = year, linewidth = year))+
geom_point()

#5
penguins %>%
ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species))+
geom_point()

penguins %>%
ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species))+
geom_point()+
facet_wrap(~species)

#6

ggplot(
  data = penguins,
  mapping = aes(
    x = bill_length_mm, y = bill_depth_mm, 
    color = species, shape = species
  )
) +
  geom_point() +
  labs(color = "Species")

ggplot(
  data = penguins,
  mapping = aes(
    x = bill_length_mm, y = bill_depth_mm, 
    color = species, shape = species
  )
) +
  geom_point() +
  labs(color = "Species", shape = "Species")
  
#7
#They show island or species on x, and fill by the other variable - thereby comparing island or species, depending on what is on X
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")
ggplot(penguins, aes(x = species, fill = island)) +
  geom_bar(position = "fill")

#2.6.1

#1
#The last plot is always saved

#2
#you change file ending to .pdf
#you can check available formats using ?ggsave()