---
title: "Chapter 13 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(nycflights13)
```

### 14

---

1.

```{r}

flights |> count(is.na(dep_time))

```

2.

```{r}

count_dest <- flights |> count(dest, sort = TRUE)

count_tailnum <- flights |> count(tailnum, wt = distance)

summarise_dest <- flights |> 
  summarise(n = n(), .by = dest) |> 
  arrange(desc(n))

summarise_tailnum <- flights |> 
  summarise(n = sum(distance), .by = tailnum) |> 
  arrange(tailnum)

```

### 14.4.8

1.

```{r, eval = TRUE}

flights |> #Work against flights data
  group_by(hour = sched_dep_time %/% 100) |> #group by sched_dep_time after integer division by 100
  summarize(prop_cancelled = mean(is.na(dep_time)), n = n()) |> #calculate proportion of canceled flights (dep_time == NA) and number of flights
  filter(hour > 1) |> #filter out the one flight at 01.00
  ggplot(aes(x = hour, y = prop_cancelled)) + #call ggplot and set x and y
  geom_line(color = "grey50") + # draw a grey line
  geom_point(aes(size = n)) #draw a point, size determined by number of flights at a given hour

```

2.

Radians, not degrees. See: https://stat.ethz.ch/R-manual/R-devel/library/base/html/Trig.html

3.

```{r}

flights_fractional_hour <- flights |>
  filter(month == 1, day == 1) |> 
  select(sched_dep_time, dep_time, dep_delay) |> 
  mutate(
    sched_dep_time_fraction = sched_dep_time %/% 100 + sched_dep_time %% 100 / 60,
    dep_time_fraction = dep_time %/% 100 + dep_time %% 100 / 60
  ) 

flights_fractional_hour

  ggplot(
    flights_fractional_hour,
    aes(sched_dep_time_fraction, dep_time_fraction)) +
  geom_point()


```

4.

```{r}

flights |> 
  select(dep_time, arr_time) |> 
  mutate(
    dep_time_5m = round(dep_time / 5) * 5,
    arr_time_5m = round(arr_time / 5) * 5
  )

```

### 14.5.4

1.

```{r}

flights |> 
  mutate(arr_delay_rank = dense_rank(desc(arr_delay))) |> 
  relocate(arr_delay_rank, arr_delay, .before = 1) |> 
  arrange(arr_delay_rank)

```

2.

```{r}

flights |> 
  group_by(tailnum) |> 
  mutate(
    mean_delayed_tailnum = mean(arr_delay, na.rm = TRUE),
    .keep = "used"
  ) |> 
  arrange(desc(mean_delayed_tailnum)) |> 
  select(!arr_delay)


```

3.

Morning

4.

`flights |> group_by(dest) |> filter(row_number() < 4)` returns up to three  
rows per group.

`flights |> group_by(dest) |> filter(row_number(dep_delay) < 4)` returns the  
rows with the smallest `dep_delay`, up to three rows per group. Since the  
`dest == "LGA"` only has one observation where `dep_delay = NA`, this result in
an `NA` in the filter call and this group being removed.

5.

```{r}

flights |> 
  group_by(dest) |> 
  mutate(total_delay = sum(arr_delay, na.rm = TRUE),
         flight_delay_proportion = dep_delay / total_delay) |> 
  relocate(total_delay, flight_delay_proportion, .before = 1)

```

6.

```{r}

flights |> 
  mutate(hour = dep_time %/% 100) |> 
  group_by(year, month, day, hour) |> 
  summarize(
    dep_delay = mean(dep_delay, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |> 
  filter(n > 5) |> 
  mutate(hour = factor(hour, levels = c(1:23,0))) |> 
  ggplot(aes(x = hour, y = dep_delay, group = interaction(month, day))) +
  geom_line()

```

7.

```{r}

flights_rel_airtime <- flights |> 
  group_by(dest) |> 
  mutate(rel_airtime = air_time / mean(air_time, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(
    speed = cut(
      rel_airtime,
      breaks = c(-Inf, 0.65, 0.85, 1.25, 1.75, Inf),
      labels = c("suspiciously fast", "fast", "normal", "slow", "heavy delay")
    )
  ) |> 
  arrange(rel_airtime)

flights_rel_airtime$rel_airtime |> max(na.rm = TRUE) #find max
flights_rel_airtime$rel_airtime |> min(na.rm = TRUE) #find min

flights_rel_airtime |> 
  ggplot(aes(x = rel_airtime)) +
  geom_histogram(binwidth = 0.05)


flights_rel_airtime |> 
  drop_na(rel_airtime) |> 
  filter(row_number() %in% 1:5 | 
        row_number() %in% (nrow(flights_rel_airtime |> drop_na(rel_airtime)) - 5):nrow(flights_rel_airtime |> drop_na(rel_airtime))
  )


```

8.

```{r}

#Find destinations with at least two carriers
dest_at_least_2_carrier <- flights |> 
  group_by(dest) |> 
  summarise(count = n_distinct(carrier)) |> 
  filter(count >= 2)

#Filter flights data for these destinations
flights_filtered <- flights |> 
  select(dest, carrier, dep_delay) |> 
  filter(dest %in% dest_at_least_2_carrier$dest)

#Create carrier ranks by comparing the delay by destination
flights_filtered |> 
  group_by(dest, carrier) |> #Caclulate by carriers and destination
  mutate(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) |> #Average delay
  group_by(dest) |>  #Drop carrier grouping to compare by destination
  mutate(rank_dest = dense_rank(avg_dep_delay)) |> 
  arrange(dest, avg_dep_delay) |> 
  distinct(dest, carrier, avg_dep_delay, rank_dest) |> #remove dep_delay column
  group_by(carrier) |> 
  mutate(rank_average = mean(rank_dest)) |> #Calculate the average rank by carrier
  ungroup() |> 
  mutate(rank_overall = dense_rank(rank_average)) |> #Compare the average rank between carriers
  arrange(rank_overall) |> 
  distinct(carrier, rank_overall)


```

### 14.6.7

1.


`mean` could be used when using variables from more similar observations (e.g. 
the `dep_delay` for the same carrier, airport, time of day, etc.)

`median` could instead be used when you want to avoid the influence of outliers that
might occur if you compare more broadly over the whole data.

The same reasoning can be applied to the variables `dep_delay` and `arr_delay` 
where the latter would have more underlying variables and could be suspected 
to have a larger spread.

2.


3.

