---
title: "Chapter 12 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(scales)
library(ggrepel)
library(patchwork)
library(beeswarm)
library(ggbeeswarm)

```

### 12.2.1

---

1.

```{r}

ggplot(mpg, aes(x = as_factor(year), y = hwy)) +
  geom_beeswarm()+
  facet_wrap(~ class, scales = "free_y") +
  stat_summary(fun = "mean", geom = "crossbar", color = "red", width = 0.5) +
  labs(
    x = "Year",
    y = "Highway fuel economy (mpg)",
    color = "Car type",
    title = "Fuel efficiency was not significantly improved during the last 10 years",
    subtitle = "Comparing car type, not necessarily same model",
    caption = "Data from fueleconomy.gov"
  )

```

2.

```{r}

mpg |> 
  ggplot(
    aes(
      cty,
      hwy,
      color = drv,
      shape = drv
    )
  ) +
  geom_point() +
  labs( 
    y = "Highway MPG",
    x = "City MPG",
    color = "Type of drive train",
    shape = "Type of drive train"
  )

```


### 12.3.1

1.

```{r}

mpg |> 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  geom_text(aes(x = -Inf, y = Inf, label = year)) +
  geom_text(aes(x = -Inf, y = -Inf, label = year)) +
  geom_text(aes(x = Inf, y = Inf, label = year)) +
  geom_text(aes(x = Inf, y = -Inf, label = year))

```

2.

```{r}

mpg |> 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  geom_text(aes(x = -Inf, y = Inf, label = year)) +
  geom_text(aes(x = -Inf, y = -Inf, label = year)) +
  geom_text(aes(x = Inf, y = Inf, label = year)) +
  geom_text(aes(x = Inf, y = -Inf, label = year)) +
  annotate(
    geom = "point",
    x = 4.25,
    y = 28,
    size = 5,
    color = "red",
    shape = 21
  )

```

3.

By adding a label that is unique to each facet

```{r}

mpg |> 
  mutate(
    facet = case_when(
      drv == "f" ~ "front",
      drv == "4" ~ "4 wheel",
      drv == "r" ~ "rear"
    )
  ) |> 
  ggplot(
    aes(
      x = displ,
      y = hwy
    )
  ) +
  geom_point() +
  facet_wrap( ~ drv) +
  geom_text(
    aes(
      x = 4,
      y = 25,
      label = facet
    )
  )

```

4.

`fill` controls background color.

5.  


  - `angle`: controls width of arrow head (smaller number is pointier)
  - `length`: length of arrow head (tip to base)
  - `ends`: `"last"`, `"first"`, or `"both"` - controls where to draw arrow head
  - `type`: `"open"` or `"closed"`, type of arrowhead.

```{r}


mpg |> 
ggplot(
    aes(
      x = displ,
      y = hwy
    )
  ) +
  geom_point() +
  annotate(
    data = mpg_arrows,
    geom = "segment",
    x = 4,
    y = 30,
    xend = 6,
    yend = 20,
    color = "red",
    arrow = arrow(
      angle = 50,
      length =  unit(0.25, "cm"),
      ends = "last",
      type = "open"
    )
  )

mpg |> 
ggplot(
    aes(
      x = displ,
      y = hwy
    )
  ) +
  geom_point() +
  annotate(
    data = mpg_arrows,
    geom = "segment",
    x = 4,
    y = 30,
    xend = 6,
    yend = 20,
    color = "red",
    arrow = arrow(
      angle = 50,
      length =  unit(0.25, "cm"),
      ends = "both",
      type = "closed"
    )
  )

mpg |> 
ggplot(
    aes(
      x = displ,
      y = hwy
    )
  ) +
  geom_point() +
  annotate(
    data = mpg_arrows,
    geom = "segment",
    x = 4,
    y = 30,
    xend = 6,
    yend = 20,
    color = "red",
    arrow = arrow(
      angle = 90,
      length =  unit(1, "cm"),
      ends = "first",
      type = "closed"
    )
  )


```

### 12.4.6

1.

`scale_color_gradient` needs to be replaced with `scale_fill_gradient`
```{r}

df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_color_gradient(low = "white", high = "red") +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_fixed()
```

2.  


  - `labs` takes a list of new name-value pairs, where the name should be an
aesthetic (e.g. color).
  - The different scales take arguments that are passed on to `continuous_scale`
  
3.

```{r}
presidential |>
  mutate(id = 33 + row_number()) |>
  ggplot(aes(x = start, y = id, color = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_color_manual(values = c(Republican = "#E81B23", Democratic = "#00AEF3"))
```

a. Combining the two variants that customize colors and x axis breaks.

```{r}
presidential |>
  mutate(id = 33 + row_number()) |>
  ggplot(aes(x = start, y = id, color = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_color_manual(values = c(Republican = "#E81B23", Democratic = "#00AEF3"))
```