---
title: "Chapter 17 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE}
library(tidyverse)
library(babynames)
library(ggrepel)
```

### 17.3.1

---

1.

```{r}

#Default bar plot showing counts of factor levels: data is disordered
gss_cat |> 
  ggplot() +
  geom_bar(aes(x = rincome)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Re-ordering rincome factor levels by count improves, but still confusing as 
#the rincome levels are not ordered
gss_cat |> 
  count(rincome) |> 
  mutate(rincome = fct_reorder(rincome, n)) |> 
  ggplot() +
  geom_col(aes(x = rincome, y = n)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

2.

```{r}

#Protestant is the most common relig
gss_cat |> 
  count(relig) |> 
  arrange(desc(n))

#Independent is the most common partyid
gss_cat |> 
  count(partyid, sort = TRUE) 

```

3.

```{r}

#Table view
gss_cat |>
  group_by(relig) |> 
  count(denom)

gss_cat |> 
  ggplot() +
  geom_bar(aes(x = relig, fill = denom)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

### 17.4.1

---

1.

`tvhours` does not look like it is normally distributed, mean is therefore not
informative. 

```{r}

gss_cat |> 
  ggplot() +
  geom_histogram(aes(x = tvhours))

```

2.

Let's create a bar graph of all columns that are factors, with counts for each
factor level for additional information. To my eye, everything seems arbitrary 
except for income.

```{r, fig.height= 25}

gss_cat |> 
  select(where(is.factor)) |> #Select out columns that are factors
  pivot_longer(cols = marital:denom) |> #Pivot all columns for graphing
  group_by(name, value) |> 
  count() |> #This step could be omitted if geom_bar was used
  ggplot() +
  geom_col(aes(x = n, y = value)) +
  facet_wrap(~ name, scales = "free_y", ncol = 1)

```

3.

It follows the order of factor levels, use e.g. `desc` to get the opposite.

```{r}

rincome <- gss_cat |> 
  pull(rincome) |> 
  levels()

rincome_releved <- gss_cat |> 
  mutate(
   rincome = fct_relevel(rincome, "Not applicable")
  ) |> 
  pull(rincome) |> 
  levels()

tibble(
 rincome,
 rincome_releved
)  


```

### 17.5.1

---

1.

```{r}

gss_cat |> 
  pull(partyid) |> 
  levels()

gss_cat |> 
  mutate(
    partyid = 
      fct_collapse(
        partyid,
        "Indep" = c("Ind,near rep", "Independent", "Ind,near dem"),
        "Dem" = c("Not str democrat", "Strong democrat"),
        "Rep" = c("Strong republican", "Not str republican"),
        "Other" = c("No answer", "Don't know", "Other party")
      )
  ) |> 
  group_by(year, partyid) |> 
  summarise(
    n = n(),
    .groups = "drop"
  ) |> 
  mutate(
    prop = n / sum(n),
    .by = c("year")
  ) |> 
  ggplot(aes(year, prop, color = fct_reorder2(partyid, year, prop))) +
  geom_line() +
  geom_point()



```

2.

One quick way is to lump the smaller groups by using `fct_lump_n`

```{r}

gss_cat |> 
  count(rincome)

gss_cat |> 
  mutate(
    rincome = fct_lump_n(rincome, 5)
  )|> 
  count(rincome)


```

