---
title: "Chapter 10 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)
```

### 10.2.1

---

1.

```{r}
mpg |> 
  ggplot() + 
  geom_point(aes(x = hwy, y = displ), shape = 24, fill = "pink")
```

2.

The `color = "blue"` needs to be moved out of aes when it is set to a value 
instead of mapped to a variable.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, color = "blue"))

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy), color = "blue")
```

3.

Stroke sets the width of the border (if your shape has a border).

4.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, color = displ < 5))
```

### 10.3.1

---

1. 

`geom_line()` for a line, `geom_boxplot()` for a boxplot, `geom_area()` for an 
area plot.

2.

`show.legend = FALSE` hides the geom from the legend. This was used previously
to maintain the regular plot shape to ease comparison (otherwise legend shrinks
the plot area).

3.

`se` plots the confidence interval of `geom_smooth()`.

4.

Going per row, from left to right:

```{r}
mpg |> #1
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(se = FALSE)

mpg |> #2
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(aes(group = drv), se = FALSE)

mpg |> #3
  ggplot(aes(x = displ, y = hwy, group = drv, color = drv)) + 
  geom_point() +
  geom_smooth(se = FALSE)

mpg |> #4
  ggplot(aes(x = displ, y = hwy)) + 
  geom_point(aes(group = drv, color = drv)) +
  geom_smooth(se = FALSE)

mpg |> #5
  ggplot(aes(x = displ, y = hwy, group = drv, linetype = drv)) + 
  geom_point(aes(color = drv)) +
  geom_smooth(se = FALSE)

mpg |> #6
  ggplot(aes(x = displ, y = hwy, group = drv, linetype = drv)) + 
  geom_point(aes(fill = drv), shape = 21, color = "white", stroke = 2 )

```

### 10.4.1

---

1. 

```{r}

mpg |> 
  ggplot(aes(x = hwy, y = displ)) +
  geom_point() +
  facet_wrap(~cty)

```

2.

It means that the specific `drv` and `cyl` combination is not existing in the 
data.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = drv, y = cyl))
```

3.

It controls whether the variable will be displayed side-by-side or stacked.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4.

For larger data sets, color gets busy and, unless specified, is only recommeded 
for six separate group members (colors).

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

5.

`ncol`or `nrow` controls `facet_wrap()` plot layout, and is inteded for use with
a single variable with many levels. Instead, `facet_grid()` is intended for 
facetting two separate variables against eachother in a matrix layout.

6.

It is much easier to compare when they are stacked.

```{r}
ggplot(mpg, aes(x = displ)) + 
  geom_histogram() + 
  facet_grid(drv ~ .)

ggplot(mpg, aes(x = displ)) + 
  geom_histogram() +
  facet_grid(. ~ drv)
```

7.

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~drv, ncol = 1, strip.position = "right")
```

### 10.5.1

---

1.

```{r}

diamonds |>
  summarise(
    min = min(depth),
    max = max(depth),
    median = median(depth),
    .by = cut
  ) |> 
  ggplot(aes(x = cut, y = median)) +
  geom_pointrange(aes(ymin = min, ymax = max))

```

2.

`geom_col` uses `stat_identity` instead of `stat_count` (default for `geom_bar`),
it is therefore advisable to use `geom_col` when you want to plot the values.

3.

Summary of geoms listed in the [ggplot2 reference:](https://ggplot2.tidyverse.org/reference/), pairs share the same arguments:

```{r}

geoms <- tribble(
  ~geom, ~stat, ~shape, ~use,
  "geom_abline", NA, "line", "Diagonal reference line",
  "geom_hline", NA, "line", "Horizontal reference line",
  "geom_vline", NA, "line", "Vertical reference line",
  "geom_bar", "stat_count", "bar", "Bar chart",
  "geom_col", NA, "bar", "Bar chart",
  "geom_bin_2d", "stat_bin_2d", "bin", "Heatmap of 2d bin counts",
  "geom_blank", NA , NA, "empty space",
  "geom_boxplot", "stat_boxplot", "boxplot", "A box and whiskers plot (in the style of Tukey)",
  "geom_contour", "stat_contour", "lines", "2D contours of a 3D surface",
  "geom_contour_filled", "stat_contour_filled", "line", "2D contours of a 3D surface",
  "geom_count", "stat_sum", "point", "Count overlapping points",
  "geom_density", "stat_density", "line", "Smoothed density estimates",
  "geom_density_2d", "stat_density_2d", "contour", "Contours of a 2D density estimate",
  "geom_density_2d_filled", "stat_density_2d_filled", "contour", "Contours of a 2D density estimate",
  "geom_dotplot", NA, "point", "Dot plot",
  "geom_errorbarh", NA, "errorbar", "Horizontal error bars",
  "geom_function", "stat_function", "line", "Draw a function as a continuous curve",
  "geom_hex", "stat_bin_hex", "hexagon", "Hexagonal heatmap of 2d bin counts",
  "geom_freqpoly", "stat_bin", "line", "Histograms and frequency polygons",
  "geom_histogram", "stat_bin", "bin", "Histograms and frequency polygons",
  "geom_jitter", NA, "point", "Jittered points",
  "geom_crossbar", NA, "crossbar", "Vertical intervals: lines, crossbars & errorbars",
  "geom_errorbar", NA, "errorbar", "Vertical intervals: lines, crossbars & errorbars",
  "geom_linerange", NA, "line", "Vertical intervals: lines, crossbars & errorbars",
  "geom_pointrange", NA, "line and point", "Vertical intervals: lines, crossbars & errorbars",
  "geom_map", NA, "line", "Polygons from a reference map",
  "geom_path", NA, "line", "Connect observations",
  "geom_line", NA, "line", "Connect observations",
  "geom_step", NA, "line", "Connect observations",
  "geom_point", NA, "point", "Points",
  "geom_polygon", NA, "polygon", "Polygons",
  "geom_qqline", "stat_qqline", "line", "A quantile-quantile plot",
  "geom_qq", "stat_qq", "point", "A quantile-quantile plot",
  "geom_quantile", "stat_quantile", "line", "Quantile regression",
  "geom_ribbon", "stat_align", "area", "Ribbons and area plots",
  "geom_area", "stat_align", "area", "Ribbons and area plots",
  "geom_rug", NA, "line", "Rug plots in the margins",
  "geom_segment", NA, "line", "Line segments and curves",
  "geom_line", NA, "line", "Line segments and curves",
  "geom_smooth", "stat_smooth", "line", "Smoothed conditional means",
  "geom_spoke", NA, "line", "Line segments parameterised by location, direction and distance",
  "geom_label", NA, "text", "Text",
  "geom_text", NA, "text", "Text",
  "geom_raster", NA, "raster", "Rectangles",
  "geom_rect", NA, "raster", "Rectangles",
  "geom_tile", NA, "raster", "Rectangles",
  "geom_violin", "stat_ydensity", "violin", "Violin plot",
  "geom_sf", "stat_sf", "simple feature", "Visualise sf objects"
  
        )

 geoms

```

4.

`stat_smooth` calculates the following variables (x or y, depending on orientation):

- `after_stat(x)`, the predicted value.
- `after_stat(xmin)`, lower pointwise confidence interval around the mean.
- `after_stat(xmax)`, upper pointwise confidence interval around the mean.
- `after_stat(se)`, standard error.

5. 

The difference can be seen if plotted:

```{r}

ggplot(diamonds, aes(x = cut, y = after_stat(prop))) + 
  geom_bar()

ggplot(diamonds, aes(x = cut, fill = color, y = after_stat(prop))) + 
  geom_bar()

ggplot(diamonds, aes(x = cut, y = after_stat(prop), group = 1)) + 
  geom_bar()

ggplot(diamonds, aes(x = cut, fill = color, y = after_stat(prop), group = 1)) + 
  geom_bar()

```

After setting `group` to 1, the proportion is calculated across all the different cuts - as opossed to running proportion for each cut.

### 10.6.1

---

1. 

Comparing setting different positions:

```{r}

ggplot(mpg, aes(x = cty, y = hwy)) + 
  geom_point()

ggplot(mpg, aes(x = cty, y = hwy)) + 
  geom_point(position = "identity")

ggplot(mpg, aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")

```

2.

Nothing, setting it to identity produces the same plot:

```{r}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(position = "identity")


```

3.

`width` and `height`, if omitted it amounts to 40% in both positive and negative direction (i.e. total spread is twice).

4.

```{r}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_count()

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_jitter()


```

5.

`position = dodge2`:

```{r}

ggplot(mpg, aes(x = drv, y = hwy, fill = model)) +
  geom_boxplot()

```

