---
title: "Chapter 3 excercises"
author: "Hugo Åkerstrand"  
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)  
library(nycflights13)
```




### 4.2.5

---
1. 

```{r}

flights |>
  filter(arr_delay > 120) #Filter out flights that arrive more than two hours delayed

flights |>
  filter(dest %in% c("IAH", "HOU"))  #Filter out flights with destination Huoston

#Create a vector to filter flights by American, United, Delta
carrier_codes <- airlines |> filter(grepl("American|United|Delta",airlines$name)) |>
  select(carrier) |>
  unlist(use.names = F)

flights |>
  filter(carrier %in% carrier_codes) #Filter out flights by vector

flights |>
  filter(month %in% c(7, 8, 9)) #Filter out flights that happened in July, August, & September

flights |>
  filter(dep_delay == 0 & arr_delay > 120) #Filter out flights that departed on time, but arrived more than two hours delayed

flights |>
  filter(arr_delay >= 60 & (sched_arr_time - sched_dep_time) -  (arr_time - dep_time) > 30) #Filter out flights that were delayed by at least an hour, but made up over 30 minutes in flight
  
```

2.

```{r}

flights |>
  arrange(desc(dep_delay), dep_time)

```

3.

```{r}

flights |> 
  arrange(desc(distance/air_time)) #Sorted on fastest flight

```

4.

```{r}

flights |>
  distinct(day,month,year) |> #Filter out each unique date entry to count the number of unique days with flight data
  count()
```

5.

```{r}
flights |>
  arrange(distance) |> #Top entry is shortest trip
  arrange(desc(distance)) #Top entry is longest trip

```

6.

Filtering before arranging results in the minimal computation.

### 4.3.5

1.

The `dep_time` is largely linear with `sched_dep_time`, the `dep_delay` is their delta

```{r}

dep_data <- flights |>
  select(dep_time, sched_dep_time, dep_delay)

dep_data
```

2.

```{r}
flights |> glimpse()
```


`dep_time` is the 4th column (type = int), `dep_delay` is the 6th (type = dbl), 
`arr_time` is the 7th (type = int), and `arr_delay` is the 9th (type = dbl).

- Most straight forward is to pass the column names or positions, although I 
personally prefer names over numbers (for clarity).
- Second most straight forward (and less verbose), would be to use the 
`starts_with()` selector.
- More labourously, one could use both positive and negative selection;
first by `select(contains(c("arr_","dep_")))` and then remove the columns 
starting with `sched`.
- More complicated options and combinations are of course available, 
but this is too complicated to be recommended.

```{r}

flights |> 
  select(dep_time, dep_delay, arr_time, arr_delay)

```

```{r}

flights |> 
  select(4, 6, 7, 9)

```

```{r}

flights |>
  select(starts_with(c("arr_", "dep_")))

```


```{r}

flights |> 
 select(
   contains(
     c("arr_","dep_")) #Include the underscore, or else arr will match to carrier as well
   ) |>
  select(
    !contains("sched")
  )

```

3.

It produces the same result as if `carrier` was only mentioned once.

```{r}
flights |>
  select(carrier, carrier)
```

4.

`any_of()` is a selection helper for a character vector. As opposed to `all_of()`, 
this selector is not strict - meaning that it doesn't check to make sure that all 
of the vector elements are found.

```{r}
variables <- c("year", "month", "day", "dep_delay", "arr_delay")

flights |> select(any_of(variables))
```

5. 

It doesn't surprise me as, select helpers have default setting `.ignore.case = TRUE`.

```{r}

flights |> select(contains("TIME"))

```

6.

```{r}

flights |>
  rename(air_time_min = air_time)

```

7.

The error is due to the fact that `select` has removed all columns except 
`tailnum`, therefore `arrange` cannot find and sort based on `arr_delay`

### 4.5.7

1.

Let's first look at what carrier has the worst average departure delays (using `dep_delay`)

```{r}

origin_delay <- flights |> 
  summarise(
    n = n(),
    delays = mean(dep_delay, na.rm = T), #calculate the average delay
    .by = carrier #group by dest
  ) |> 
  arrange(desc(delays))

origin_delay |>
  ggplot()+
  geom_col(aes(x = fct_reorder(carrier, delays), y = delays))

```

However, a lot of the data represent `carrier` and `dest`combinations with few
observations:

```{r}
flights |> 
  group_by(carrier, dest) |> 
  summarize(n = n(), .groups = "drop") |> 
  ggplot()+
  geom_histogram(aes(x=n), binwidth = 100)
```

Let's filter the `carrier` and `dest` flights with less than 100 observations, 
and then plot the `mean(dep_delay)` for a given `carrier` and `dest`:

```{r}
flights |> 
  mutate(n = n(),
         .before = 1,
         .by = c(carrier, dest)) |> 
  filter(n >= 100) |> 
  summarize(n = n(), 
            delays = mean(dep_delay, na.rm = T), 
            .by = c(carrier,dest)
            ) |> 
  ggplot()+
  geom_point(aes(x = dest, y = delays, color = carrier),
             position = position_dodge(.9))+
  facet_wrap( ~ carrier)

origin_delay
```

By looking at these two plots, we can see the difficulty in comparing the average 
delay per carrier in the first graph - some carriers only have one specific flight 
and become very sensitive to this specific route (e.g. `F9`). At the same time, 
comparing DL to EV shows that clearly the latter is having a systematic issue with
delays across all its destinations.

2.

```{r}

flights |> 
  slice_max(dep_delay, n = 1, by = dest)
  

```

3.

Delays are much worse in the evening

```{r}

flights |> 
  summarise(
    delays = mean(dep_delay,
                  na.rm = TRUE),
    .by = hour
  ) |> 
  ggplot()+
  geom_col(aes(x = hour, y = delays))

```

4.

A negative value passed to `n` will make `slice_min()` ignore that many rows of the group - so -1 will make a group of 5 only use 4 rows.

5.

`count()` will calculate the total observations of the group. 
If you specify `sort = TRUE` it will put the most numerous observation on the 
top of the output.

6.

a. `group_by(y)` will result in two groups belonging either to a or b.

```{r}
df <- tibble(
  x = 1:5,
  y = c("a", "b", "a", "a", "b"),
  z = c("K", "K", "L", "L", "K")
)

df |>
  group_by(y)
```

b. It will arrange `df` according to `df$y` (a, then b) - the df is not grouped!

```{r}
df |> 
  arrange(y)
```

c. It will give give a 2x2 tibble: a `y` column (a or b) and `mean_x` column of x

```{r}
df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
```

d. Same as above, but a third column `z` and another row as a consequence of 
`group_by(y, z)`. The message states that the output is now grouped by only `y`,
rather than also `z`.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))
```

e. It is different in not maintaining any grouping variable.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop")
```

f. They are different in `mutate` adding a column, and `summarize` producing a
new one, consisting of grouping variables and the `mean_x`.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  mutate(mean_x = mean(x))
```

